PubCrawl: A Publication Search Engine

### 2012-05-17 report/paper.tex
Expand important details about Lucene
Talk about what Lucene scoring is doing under the covers and how it affects
what we're trying to do

We use Apache Lucene \cite{lucene} to index documents and provide scores based
on the content of documents. We build the usual inverted index over the corpus,
filtering out stopwords. The structure of the documents is preserved in the XML
corpus, so we can index titles, subtitles, abstracts, and full text as separate
fields. We also include various metadata to simplify the implementation of the
front-end.

Lucene uses the Boolean model \cite{lancaster73} to filter documents and the
Vector Space Model \cite{salton75} to score documents. Documents and queries
are vectors with each term a dimension and coordinates equal to tf*idf values.
Then, the relevance of a document to a query is related to the angle between
them (the cosine similarity score). Lucene generalizes the cosine similarity
score in several ways. First, documents are allowed to have multiple and
duplicate fields. Second, Lucene allows index-time boosting of documents and
fields as well as query-time boosting of query terms. Finally, Lucene does not
require every term to appear in hits, instead using a coordination factor to
score documents which contain more terms higher. These features can be
exploited to improve the quality of search results.

>>>>

We use Apache Lucene \cite{lucene} to index documents and compute relevance
scores. We build the usual inverted index over the corpus, filtering out
stopwords. The structure of the documents is preserved in the XML corpus, so we
can easily parse titles, subtitles, abstracts, and full text. Lucene supports
multi-field documents, so we index and perform searches on each independently.
We also include various metadata to simplify the implementation of the
front-end.

Lucene uses the Boolean model \cite{lancaster73} to filter documents and the
Vector Space Model \cite{salton75} to score documents. Documents and queries
are vectors with each term a dimension and coordinates equal to tf*idf values.
Then, the relevance of a document to a query is related to the angle between
them (the cosine similarity score). Lucene generalizes the cosine similarity
score in several ways. First, documents are allowed to have multiple and
duplicate fields. Second, Lucene allows index-time boosting of documents and
fields as well as query-time boosting of query terms. Finally, Lucene does not
require every term to appear in hits, instead using a coordination factor to
score documents which contain more terms higher.

While these features can be exploited to improve the quality of search results,
they also have some unintended side effects. In particular, Lucene's scoring is
biased towards matches in shorter fields due to the way it normalizes even
though fields are boosted equally by default. We investigated using user
inputted weights to boost individual fields within Lucene's scoring mechanism,
but found it as difficult to combine the resulting scores with PageRank as
performing three separate searches and combining four scores.
